---
title: "SoccerWorksheetKey"
format: docx
---

```{r}
# Load packages
library(readr)
library(dplyr)
library(ggplot2)
library(GGally)
library(tidyr)
library(cluster)
library(factoextra)
library(fmsb)

# Load cleaned data
fifa_clean <- read_csv("fifa_clean.csv", show_col_types = FALSE)
```

# ---- PART 1: EXPLORATORY ANALYSIS ----

# Q1: What are the average values for age, overall rating, and key skills?

```{r}
summary_stats <- fifa_clean |>
  summarise(
    avg_age = mean(age),
    avg_overall = mean(overall),
    avg_pace = mean(pace),
    avg_shooting = mean(shooting),
    avg_passing = mean(passing),
    avg_dribbling = mean(dribbling),
    avg_defending = mean(defending),
    avg_physic = mean(physic)
  )
summary_stats  # Answer: Shows averages (e.g., avg_overall ≈ 66.4)
```

# Q2: What does the distribution of overall player ratings look like?

```{r}
ggplot(fifa_clean, aes(x = overall)) +
  geom_histogram(bins = 30, fill = "pink", color = "white") +
  labs(title = "Distribution of Overall Ratings", x = "Overall", y = "Count") +
  theme_minimal()
# Answer: Most players have overall ratings between 60–75, with few outliers above 85.
```

# Q3: How does overall rating relate to potential?

```{r}
ggplot(fifa_clean, aes(x = overall, y = potential)) +
  geom_point(alpha = 0.3, color = "deeppink") +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  labs(title = "Overall vs Potential", x = "Overall", y = "Potential") +
  theme_minimal()
# Answer: There is a strong positive linear relationship — better current players often have higher potential.
```

# Q4: Create a pairwise correlation plot of all skill attributes.

```{r}
skill_subset <- fifa_clean |>
  select(overall, potential, pace, shooting, passing, dribbling, defending, physic)

ggpairs(skill_subset, lower = list(continuous = wrap("smooth", alpha = 0.3, size = 0.1)),
        title = "Pairwise Skill Attribute Correlations")
# Answer: Overall rating is most correlated with physic, dribbling, and shooting. Passing is also moderately correlated.
```

# ---- PART 2: K-MEANS CLUSTERING ----

```{r}

fifa_cluster <- fifa_clean |> select(pace, shooting, passing, dribbling, defending, physic)
fifa_scaled <- scale(fifa_cluster)
set.seed(123)
k4 <- kmeans(fifa_scaled, centers = 4, nstart = 20)
fifa_clean <- fifa_clean |> mutate(Cluster = as.factor(k4$cluster))
```

# Q5: Use K-means to group players into 4 clusters — done above.

# Q6: Plot the standardized skill profiles for each cluster

```{r}
centers <- as_tibble(k4$centers) |> mutate(Cluster = as.factor(1:4)) |> 
  pivot_longer(-Cluster, names_to = "Attribute", values_to = "Value")

ggplot(centers, aes(x = Attribute, y = Value, fill = Cluster)) +
  geom_col(position = "dodge") +
  labs(title = "Cluster Skill Profiles", y = "Standardized Skill Value") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
# Answer: Each cluster emphasizes different skills — e.g., Cluster 1 may be strong defenders while Cluster 3 may be attackers.
```

# Q7: Visualize clusters using PCA

```{r}
fviz_cluster(k4, data = fifa_scaled, geom = "point", palette = "jco", ggtheme = theme_minimal())
# Answer: Players form 4 distinguishable groups based on their skill attributes.
```

# ---- PART 3: CLUSTER INTERPRETATION ----

# Q8: How do average raw skill ratings vary across clusters?

```{r}
cluster_summary <- fifa_clean |>
  group_by(Cluster) |>
  summarise(across(c(pace, shooting, passing, dribbling, defending, physic), mean))
cluster_summary
# Answer: Each cluster specializes — some have higher dribbling and pace, others in defending and physic.
```

# Q9: Which cluster has the most players?

```{r}
cluster_sizes <- fifa_clean |> count(Cluster, sort = TRUE)
cluster_sizes
ggplot(cluster_sizes, aes(x = Cluster, y = n, fill = Cluster)) +
  geom_col() +
  labs(title = "Number of Players in Each Cluster", x = "Cluster", y = "Count") +
  theme_minimal()
# Answer: Cluster with most players likely represents well-rounded or average-skill players.

```

# Q10: How are clusters distributed among top 5 nationalities?

```{r}
top_countries <- fifa_clean |>
  count(nationality_name, sort = TRUE) |>
  slice_max(n, n = 5) |>
  pull(nationality_name)

fifa_top <- fifa_clean |> filter(nationality_name %in% top_countries)

ggplot(fifa_top, aes(x = nationality_name, fill = Cluster)) +
  geom_bar(position = "dodge") +
  labs(title = "Cluster Distribution by Country", x = "Country", y = "Number of Players") +
  theme_minimal()
# Answer: Distributions vary — e.g., Brazil and France may dominate attacker or technical clusters.
```

# ---- PART 4: PLAYER ARCHETYPES ----

```{r}
# Create radar chart of player archetypes by cluster

# Step 1: Compute means
radar_data <- fifa_clean |>
  group_by(Cluster) |>
  summarise(
    Overall = mean(overall),
    Pace = mean(pace),
    Shooting = mean(shooting),
    Passing = mean(passing),
    Dribbling = mean(dribbling),
    Defending = mean(defending),
    Physic = mean(physic)
  ) |> 
  as.data.frame()

# Step 2: Convert to proper format for radar chart
rownames(radar_data) <- radar_data$Cluster
radar_data <- radar_data[, -1]

# Step 3: Add max and min rows
radar_plot_data <- rbind(
  rep(100, 7),  # max
  rep(0, 7),    # min
  radar_data    # actual cluster data
)

# Step 4: Draw radar chart
radarchart(radar_plot_data, axistype = 1,
           pcol = c("pink", "hotpink", "blanchedalmond", "orchid"),
           pfcol = c("#FF69B480", "#FF149380", "#FFDAB980", "#DA70D680"),
           plwd = 2, plty = 1,
           cglcol = "grey", cglty = 1,
           axislabcol = "grey", vlcex = 0.8,
           title = "Radar Profiles of Player Clusters")

legend("topright", legend = paste("Cluster", 1:4),
       col = c("pink", "hotpink", "blanchedalmond", "orchid"),
       lty = 1, lwd = 2, bty = "n")

# Answer: Radar plots clearly show how player types differ — attackers, defenders, balanced, physical, etc.

```