---
title: "Score"
format: html
---

```{r}
library(readr)
library(tidyverse)

powerlift_raw <- read_csv("C:\Users\sarah\Downloads\openpowerlifting-latest%20%(1)\openpowerlifting-2025-03-29\openpowerlifting-2025-03-29-56319e49.csv")

library(readr)
powerlift_raw <- read_csv("C:/Users/sarah/Downloads/openpowerlifting-latest (1)/openpowerlifting-2025-03-29/openpowerlifting-2025-03-29-56319e49.csv")

powerlift <- powerlift_raw |>
  group_by(Name) |>
  arrange(desc(Date), .by_group = TRUE) |>
  slice(1)







library(here)
powerlift <- read_csv("smshel21/powerlifting/exploration/data/powerlifting_data_1.csv")

powerlift <- powerlift_raw |>
  group_by(Name) |>
  arrange(desc(Date), .by_group = TRUE) |>
  slice(1)

#write.csv(powerlift, file = "powerlifting_data_1.csv") 
```

```{r}
power <- read.csv("smshel21/data/powerlifting_data.csv")

powerlift_sample <- power |>
  ungroup() |>
  #drop_na(# list off the variables you know you want to use)
  sample_n(10000)

powerlift_sample |>
  write_csv(file = "powerlift_sample_data.csv")
```

```{r}
powerlift <- read.csv("smshel21/data/powerlift_sample_data.csv")


powerlift_new <- powerlift |>
  filter(Event == "SBD") |>
  filter(Equipment %in% c("Raw", "Single-ply")) |>
  filter(Place != "DQ") |>
  filter(Place != "NS") |>
  drop_na(Place) |>
  filter(Place != "G") |>
  filter(Place != "DD") |>
  drop_na(Age) |>
  drop_na(AgeClass)

powerlift_new |>
  group_by(AgeClass) |>
  tally()

powerlift_new |>
  write_csv(file = "powerlift_data_new.csv")

powerlift_n <- read.csv("smshel21/powerlifting/exploration/data/powerlift_data_new.csv")
  
  
  
```


# Exploratory Analysis
```{r}
ggplot(data = powerlift, aes(x = Equipment, Wilks)) +
  geom_col()
```

```{r}
ggplot(data = powerlift, aes(x = Sex, Best3BenchKg)) +
  geom_col()
```

```{r}
ggplot(data = powerlift, aes(x = BodyweightKg, y = Deadlift2Kg)) +
  geom_point()
```

```{r}

```


# Test Models

Better - test 1 or test 2?
```{r}
test1 <- lm(Wilks ~ Age + BodyweightKg + Best3BenchKg, data = powerlift)
summary(test1)
```

```{r}
test2 <- lm(Wilks ~ Age + BodyweightKg + Best3BenchKg + Best3SquatKg, data = powerlift)
summary(test2)
```

Bad predictor (cont)
```{r}
rn <- lm(Glossbrenner ~ BodyweightKg + Wilks + Best3SquatKg + Squat2Kg + Age, data = powerlift)
summary(rn)
```



t-values
```{r}
# high vs low t-value
wt <- lm(Dots ~ BodyweightKg + Wilks + Best3SquatKg, data = powerlift)
summary(wt)
```


# co
```{r}
test3 <- lm(TotalKg ~ Best3SquatKg + Best3BenchKg + Best3DeadliftKg, data = powerlift)
summary(test3)
```

```{r}
ggplot(data = powerlift, aes(x = Best3BenchKg, y = TotalKg)) +
  geom_point()
```

```{r}
total <- powerlift |>
  select(TotalKg, Best3SquatKg, Best3BenchKg, Best3DeadliftKg)

cov(total)
```


```{r}
#ggplot(data = powerlift, aes(x = Sex, y = Best3DeadliftKg)) +
 #geom_col()
```



The Powerlifting dataset features data from individual lifters during their most recent competitions. Variables include characteristics such as sex, age, birth year and age classes, and division. Other variables include 3 types of weightlifting: squats, benches, and deadlifts. 

- Wilks measures relative strength of weightlifters despite different weight classes
- Glossbrenner evaluates strength relative to bodyweight
- Dots is the ratio between bodyweight and the weight lifted
- Goodlift is the scoring system used

The goal of this module is to attempt to use different variables to predict different aspects of powerlifting or information about the lifter themselves.


```{r}
# males to start -  exploratory
# ggplot, smooothers - quadratic formula, find appropriate power
# add females and compare - use to explain difference in fits and how to choose which one out of the two
# show incorrect models, have them explain why?
# assume they know higher power and also indicator variables but not both together
# have them build models for boys and compare different power polynomials
# use calculus or R to find max value

powerlift$SexIND = ifelse(powerlift$Sex == "M", 1, 0)


ggplot(data = powerlift, aes(x = Sex, y = Deadlift3Kg)) +
  geom_boxplot()
```

Try to use Deadlift3Kg and Sex to predict Age

a) Create a plot
```{r}
ggplot(data = powerlift, aes(x = Age, y = Deadlift3Kg)) +
  geom_point()
```

b) Create a model. Are Deadlift3Kg and SexIND useful predictors?
```{r}
powerlift$SexIND = ifelse(powerlift$Sex == "M", 1, 0)

age1 <- lm(data = powerlift, Age ~ SexIND + Deadlift3Kg)
summary(age1)
```

c) Try Best3DeadliftKg in place of Deadlift3Kg.
```{r}
age2 <- lm(data = powerlift, Age ~ SexIND + Best3DeadliftKg)
summary(age2)
```

d) Write Model


2.

a)
```{r}
ggplot(data = powerlift, aes(x = BodyweightKg, y = Best3DeadliftKg)) +
  geom_point()

ggplot(data = powerlift, aes(x = BodyweightKg, y = Squat3Kg)) +
  geom_point()
```


b)
```{r}
age3 <- lm(data = powerlift, BodyweightKg ~ SexIND + Best3BenchKg)
summary(age3)
```

```{r}
age4 <- lm(data = powerlift, BodyweightKg ~ SexIND + Bench3Kg)
summary(age4)
```

c) Try more

```{r}
ggplot(data = powerlift, aes(x = BodyweightKg, y = Bench3Kg)) +
  geom_point()

ggplot(data = powerlift, aes(x = BodyweightKg, y = Bench2Kg)) +
  geom_point()
```

```{r}
age4.1 <- lm(data = powerlift, BodyweightKg ~ SexIND + Bench3Kg + Bench2Kg)
summary(age4.1)
```

```{r}
age4.2 <- lm(data = powerlift, BodyweightKg ~ SexIND + Bench3Kg + Bench2Kg + Squat3Kg)
summary(age4.2)
```


d) Model



3.

```{r}
age5 <- lm(data = powerlift, Goodlift ~ SexIND + Glossbrenner + Wilks + Dots + Best3DeadliftKg)
summary(age5)
```

```{r}
age6 <- lm(data = powerlift, Goodlift ~ SexIND + Glossbrenner)
summary(age6)
```






