---
title: "Score Worksheet Rough Draft"
author: "Sarah Sheldon"
format: 
  html:
    embed-resources: true
format: html
---

The Powerlifting dataset features data from individual lifters during their most recent competitions. Variables include characteristics such as sex, age, birth year and age classes, and division. Other variables include 3 types of weightlifting: squats, benches, and deadlifts. 

- Wilks measures relative strength of weightlifters despite different weight classes
- Glossbrenner evaluates strength relative to bodyweight
- Dots is the ratio between bodyweight and the weight lifted
- Goodlift is the scoring system used
- AgeClass: age group competed in
- Age: Age of the participant
- Best3DeadliftKg: highest weight deadlifted out of the 1kg, 2kg, and 3kg lifts (i.e. best out of 3?)
- Sex: Sex of the participant

- edits made to original data set:
  - now an equal amount of age classes (500 each)
  - got rid of non-numeric entries in the place variable
  - got rid of missing/blank values in Age and AgeClass

For this worksheet, we want to use Sex and Age as predictors for Best3DeadliftKg.

1. 
a) Create a scatterplot comparing Age and Best3DeadliftKg. Add a smoother to incorporate one sex. What do you notice about the curve? 
```{r}

# add squat and bench?
library(tidyverse)
library(readr)
powerlift <- read.csv("smshel21/powerlifting/module/data/powerlift_sample_data_updated.csv")

powerlift_n <- read.csv("smshel21/powerlifting/exploration/data/powerlift_data_new.csv")

powerlift_new <- powerlift |>
  filter(Event == "SBD") |>
  filter(Equipment %in% c("Raw", "Single-ply")) |>
  filter(Place != "DQ") |>
  filter(Place != "NS") |>
  drop_na(Place) |>
  filter(Place != "G") |>
  filter(Place != "DD") |>
  drop_na(Age) |>
  drop_na(AgeClass)


powerlift_new$SexIND = ifelse(powerlift_new$Sex == "M", 1, 0)


library(tidyverse)
#powerlift <- read.csv("smshel21/data/powerlift_sample_data_updated.csv")


#powerlift$SexIND = ifelse(powerlift$Sex == "M", 1, 0)

powerlift_m <-  powerlift_new |>

  filter(Sex == "M")

ggplot(data = powerlift_m, aes(x = Age, y = Best3DeadliftKg)) +
  geom_point() +
  geom_smooth(aes(colour = Sex))

powerlift_f <- powerlift_new |>
  filter(Sex == "F")

ggplot(data = powerlift_f, aes(x = Age, y = Best3DeadliftKg)) +
  geom_point() +
  geom_smooth(color = "blue", se = FALSE) +
  geom_smooth(color = "red", method = "lm", formula = y ~ poly(x, 5), se = FALSE)
```





b) By just looking at the graph, what type of model should be used?





c) Now, add another smoother using y ~ poly(x, 2) to compare to the male line. This adds a quadratic curve. Does it fit? Try editing the number to find the curve that best matches.
```{r}
male <- lm(data = powerlift_m, Best3DeadliftKg ~ poly(Age, 4))
predict_m <- predict(object = male, newdata = tibble(Age = powerlift_m$Age))
male_peak <- powerlift_m$Age[which.max(predict_m)]


female <- lm(data = powerlift_f, Best3DeadliftKg ~ poly(Age, 4))
predict_f <- predict(object = female, newdata = tibble(Age = powerlift_f$Age))
female_peak <- powerlift_f$Age[which.max(predict_f)]

ggplot(data = powerlift_new, aes(x = Age, y = Best3DeadliftKg)) +
  geom_point(alpha = .6) +
  geom_smooth(colour = "turquoise", se = FALSE) +
 geom_smooth(colour = "red", method = "lm", formula = y ~ poly(x, 4), se = FALSE) +
  facet_grid(Sex~., scales = "free_y") +
  guides(colour = "none") +
  geom_vline(xintercept = male_peak, colour = "purple", size = 1) +
  geom_vline(xintercept = female_peak, colour = "orange", size = 1) +
  theme_bw() +
  labs(y = "BestDeadliftKG", title = "Impact of Age on Best Deadlifted Weight Among Different Sexes")
# confidence interval
```




d) Create a model based off of the curve that best fits. What do you notice?

```{r}
DL <- lm(data = powerlift_new, Best3DeadliftKg ~ Age + SexIND + I(Age^2))
summary(DL)
```





2. Add a smoother that shows the female sex and use y ~ poly(x, 2) to see which curve best fits.
```{r}

ggplot(data = powerlift_f, aes(x = Age, y = Best3DeadliftKg)) +
  geom_point() +
  geom_smooth(aes(colour = Sex)) +
 geom_smooth(aes(colour = Sex), method = "lm", formula = y ~ poly(x, 4))

```

a. Notice that the female curve best fits a quadratic model based on the graph. However, a quadratic model is incorrect for this situation. Why is that?

The male curve fits a higher exponent, so we need to account for it to make sure that males are included in the model.



b. Create a model to predict Best3DeadliftKg.

```{r}
DL_fin <- lm(data = powerlift_new, Best3DeadliftKg ~ Age + I(Age^5) + SexIND + Age:SexIND + I(Age^5):SexIND)
summary(DL_fin)
```



```{r}
library(broom)
model <- lm(data = powerlift_new, Best3DeadliftKg ~ Age + SexIND + Age:SexIND + I(Age^2) + I(Age^2):SexIND + I(Age^3) + I(Age^4))
summary(model)

male <- lm(data = powerlift_m, Best3DeadliftKg ~ poly(Age, 4))
predict_m <- predict(object = male, newdata = tibble(Age = powerlift_m$Age))
male_peak <- powerlift_m$Age[which.max(predict_m)]


female <- lm(data = powerlift_f, Best3DeadliftKg ~ poly(Age, 4))
predict_f <- predict(object = female, newdata = tibble(Age = powerlift_f$Age))
female_peak <- powerlift_f$Age[which.max(predict_f)]

```





.

