---
title: "Score Worksheet Rough Draft"
<<<<<<< HEAD
author: "Sarah Sheldon"
format: 
  html:
    embed-resources: true
=======
format: html
>>>>>>> 6b58523b62f9e6d0357fe38e4ff9df7d4baf5747
---

The Powerlifting dataset features data from individual lifters during their most recent competitions. Variables include characteristics such as sex, age, birth year and age classes, and division. Other variables include 3 types of weightlifting: squats, benches, and deadlifts. 

- Wilks measures relative strength of weightlifters despite different weight classes
- Glossbrenner evaluates strength relative to bodyweight
- Dots is the ratio between bodyweight and the weight lifted
- Goodlift is the scoring system used
- AgeClass: age group competed in
- Age: Age of the participant
- Best3DeadliftKg: highest weight deadlifted out of the 1kg, 2kg, and 3kg lifts (i.e. best out of 3?)
- Sex: Sex of the participant

- edits made to original data set:
  - now an equal amount of age classes (500 each)
  - got rid of non-numeric entries in the place variable
  - got rid of missing/blank values in Age and AgeClass

For this worksheet, we want to use Sex and Age as predictors for Best3DeadliftKg.

1. 
a) Create a scatterplot comparing Age and Best3DeadliftKg. Add a smoother to incorporate one sex. What do you notice about the curve? 
```{r}
<<<<<<< HEAD
# add squat and bench?
library(tidyverse)
library(readr)
powerlift <- read.csv("smshel21/powerlifting/module/data/powerlift_sample_data_updated.csv")

powerlift_n <- read.csv("smshel21/powerlifting/exploration/data/powerlift_data_new.csv")

powerlift_new <- powerlift |>
  filter(Event == "SBD") |>
  filter(Equipment %in% c("Raw", "Single-ply")) |>
  filter(Place != "DQ") |>
  filter(Place != "NS") |>
  drop_na(Place) |>
  filter(Place != "G") |>
  filter(Place != "DD") |>
  drop_na(Age) |>
  drop_na(AgeClass)


powerlift_n$SexIND = ifelse(powerlift_n$Sex == "M", 1, 0)

powerlift_m <-  powerlift_n |>
=======
library(tidyverse)
powerlift <- read.csv("smshel21/data/powerlift_sample_data_updated.csv")


powerlift$SexIND = ifelse(powerlift$Sex == "M", 1, 0)

powerlift_m <-  powerlift |>
>>>>>>> 6b58523b62f9e6d0357fe38e4ff9df7d4baf5747
  filter(Sex == "M")

ggplot(data = powerlift_m, aes(x = Age, y = Best3DeadliftKg)) +
  geom_point() +
  geom_smooth(aes(colour = Sex))
```





b) By just looking at the graph, what type of model should be used?





c) Now, add another smoother using y ~ poly(x, 2) to compare to the male line. This adds a quadratic curve. Does it fit? Try editing the number to find the curve that best matches.
```{r}
ggplot(data = powerlift_m, aes(x = Age, y = Best3DeadliftKg)) +
  geom_point() +
  geom_smooth(aes(colour = Sex)) +
 geom_smooth(aes(colour = Sex), method = "lm", formula = y ~ poly(x, 5))
```




d) Create a model based off of the curve that best fits. What do you notice?
<<<<<<< HEAD

=======
```{r}
DL <- lm(data = powerlift_m, Best3DeadliftKg ~ Age + SexIND + I(Age^2) )
```
>>>>>>> 6b58523b62f9e6d0357fe38e4ff9df7d4baf5747




2. Add a smoother that shows the female sex and use y ~ poly(x, 2) to see which curve best fits.
```{r}
<<<<<<< HEAD
ggplot(data = powerlift_n, aes(x = Age, y = Best3DeadliftKg)) +
  geom_point() +
  geom_smooth(aes(colour = Sex)) +
 geom_smooth(aes(colour = Sex), method = "lm", formula = y ~ poly(x, 4))
=======
ggplot(data = powerlift, aes(x = Age, y = Best3DeadliftKg)) +
  geom_point() +
  geom_smooth(aes(colour = Sex)) +
 geom_smooth(aes(colour = Sex), method = "lm", formula = y ~ poly(x, 5))
>>>>>>> 6b58523b62f9e6d0357fe38e4ff9df7d4baf5747
```

a. Notice that the female curve best fits a quadratic model based on the graph. However, a quadratic model is incorrect for this situation. Why is that?




<<<<<<< HEAD
b. Create a model to predict Best3DeadliftKg.
=======
b. Create a model to predict Best3DeadliftKg
>>>>>>> 6b58523b62f9e6d0357fe38e4ff9df7d4baf5747






.

