---
title: "Score Worksheet Rough Draft"
author: "Sarah Sheldon"
format: 
  html:
    embed-resources: true
---

The Powerlifting dataset features data from individual lifters during their most recent competitions. Variables include characteristics such as sex, age, birth year and age classes, and division. Other variables include 3 types of weightlifting: squats, benches, and deadlifts. 

- Wilks measures relative strength of weightlifters despite different weight classes
- Glossbrenner evaluates strength relative to bodyweight
- Dots is the ratio between bodyweight and the weight lifted
- Goodlift is the scoring system used
- AgeClass: age group competed in
- Age: Age of the participant
- Best3DeadliftKg: highest weight deadlifted out of the 1kg, 2kg, and 3kg lifts (i.e. best out of 3?)
- Sex: Sex of the participant

- edits made to original data set:
  - now an equal amount of age classes (500 each)
  - got rid of non-numeric entries in the place variable
  - got rid of missing/blank values in Age and AgeClass

For this worksheet, we want to use Sex and Age as predictors for Best3DeadliftKg.

1. 
a) Create a scatterplot comparing Age and Best3DeadliftKg. Add a smoother to incorporate one sex. What do you notice about the curve? 
```{r}
library(tidyverse)
powerlift <- read.csv("smshel21/data/powerlift_sample_data_updated.csv")


powerlift$SexIND = ifelse(powerlift$Sex == "M", 1, 0)

powerlift_m <-  powerlift |>
  filter(Sex == "M")

ggplot(data = powerlift_m, aes(x = Age, y = Best3DeadliftKg)) +
  geom_point() +
  geom_smooth(aes(colour = Sex))
```





b) By just looking at the graph, what type of model should be used?





c) Now, add another smoother using y ~ poly(x, 2) to compare to the male line. This adds a quadratic curve. Does it fit? Try editing the number to find the curve that best matches.
```{r}
ggplot(data = powerlift_m, aes(x = Age, y = Best3DeadliftKg)) +
  geom_point() +
  geom_smooth(aes(colour = Sex)) +
 geom_smooth(aes(colour = Sex), method = "lm", formula = y ~ poly(x, 2))
```




d) Create a model based off of the curve that best fits. What do you notice?





2. Add a smoother that shows the female sex and use y ~ poly(x, 2) to see which curve best fits.
```{r}
ggplot(data = powerlift, aes(x = Age, y = Best3DeadliftKg)) +
  geom_point() +
  geom_smooth(aes(colour = Sex)) +
 geom_smooth(aes(colour = Sex), method = "lm", formula = y ~ poly(x, 2))
```

a. Notice that the female curve best fits a quadratic model based on the graph. However, a quadratic model is incorrect for this situation. Why is that?




b. Create a model to predict Best3DeadliftKg.






.

