---
author: "Abby Sikora"
---

```{r} 
library(shiny)
library(tidyverse)
library(readr)
library(here)
library(rvest)

url4 <- "https://herhoopstats.com/stats/ncaa/player/caitlin-clark-stats-11eb2f34-a838-c400-aa81-12df17ae4e1e/"

h4 <- read_html(url4)

tab4 <- h4 |> html_nodes("table")

clark_df <- tab4[[2]] |> html_table(fill = TRUE)

# Clean Data
clark_df$PTS <- gsub(",", "", clark_df$PTS) # Remove commas
clark_df$PTS <- as.integer(clark_df$PTS)    # Convert to integer

# Fix column names
colnames(clark_df) <- make.names(colnames(clark_df))

clark_df
```

Exploring the Data: all data lines up with a certain year of her career

Caitlin Clark college basketball full career data

season: freshman - senior year (2020-2024)

team: irrelevant

G: Games played per season

GS: irrelevant

MIN: minutes played per year

PTS: Probably focus in this -> points per season

What next?: 
- change season to freshman through senior to order data better in the shiny app
- focus on PTS, FTM vs FTA (FT%)

- Clean up app



```{r}
# Load necessary libraries
library(rvest)
library(dplyr)

# Define the URL
url <- "https://www.espn.com/womens-college-basketball/player/_/id/4433730/paige-bueckers"

# Read the webpage content
webpage <- read_html(url)

# Extract tables as a list
tables <- webpage %>% html_nodes("table") %>% html_table(fill = TRUE)

# Extract the season column from Table 2
season_table <- tables[[2]] %>% select(1)  # Selecting only the first column (Season)

# Extract the stats columns from Table 3
stats_table <- tables[[3]]  # Assuming this contains all the numeric stats

# Rename the season column for clarity
colnames(season_table) <- "Season"

# Merge the two tables by row index
beuckers_df <- cbind(season_table, stats_table)

# Print the final merged table
print(beuckers_df)

beuckers_df <- beuckers_df |>
  mutate(Season = case_when(
    Season == "2020-21" ~ "First-year",
    Season == "2021-22" ~ "Sophomore",
    Season == "2023-24" ~ "Senior",
    Season == "2024-25" ~ "5th Year",
    TRUE ~ Season
  )) |> 
  # THEN Select only the desired columns
  select(Season, PTS, `FT%`, `FG%`, `3P%`)
# Fix column names
colnames(beuckers_df) <- make.names(colnames(beuckers_df))

```

Which variables would be cool to focud on? 

Missed an entire year bc of torn ACL, took 5th year

These stats are season averages

Modify Clark Data






